DNSKEY Management

### 2012-12-06 paper/paper.tex
new figure.
Fix wrong reference name
Add some comments, some grammatical corrections, spellcheck

\twocolumn[
\maketitle
\begin{onecolabstract}
  In this paper we recapitulate the current deployment status of
  DNSSEC and go into the main issues when managing its keys. It
  outlines the paper ``Interadministrative Challenges in Managing
  DNSKEYs'' \cite{Osterweil09} and updates some of its contents
  especially with respect to the signing of the root zone on July
  15, 2012. We will see that DNSSEC penetration in the Internet's
  zones is still low but that the necessary foundations have been lain
  for full deployment.
\end{onecolabstract}
\vspace{\baselineskip}
]

@@

In this paper we will first remind of the concepts and operations in
DNSSEC relevant for key management. Then we will give the reader an
overview of the current status of DNSSEC deployment in the Internet to
put the in-detail discussion about key management into the right
context.

Finally, we will shortly outline the DNSSEC situation in Switzerland
(i.e. the \verb|ch.| zone and how its authorities handle DNSKEY management
issues.

@@

By signing the keys of the TLD's, the root will put its trust into their keys, and any keys the TLD's have signed, resulting in a trust hierarchy illustrated in fig.~\ref{fig:trust-chain}. This way, a resolver querying the LACAL's DNS server for \verb|lacal.epfl.ch.| will be able to verify the received query answer against the known root-authority in the following way:

\begin{enumerate}
\item Verify the reply's authenticity using \verb|lacal.epfl.ch.|'s public key, obtained from \verb|epfl.ch.|'s DNS server ;
\item Verify the query -- containing \verb|lacal.epfl.ch.|'s public key -- returned by \verb|epfl.ch.| with its public key, retrieved from the \verb|ch.| DNS server;
\item Again, verify \verb|ch.|'s responses by getting its public key from the root domain.
\item Finally, the root domain's responses can be verified with its already known public key.
\end{enumerate}

DNSSEC's trust hierarchy is similar to the SSL Certificate infrastructure, where an entity marks its trust in another by \emph{signing} its certificate\footnote{Which is nothing more than a public key with additional metadata.}. However, DNSSEC theoretically only has one single top-authority\footnote{The use of TAR's should remain temporary, as in an ideal setup DNSSEC could work without them.}, while the SSL Certificate hierarchy disposes of many.

\begin{figure*}
\center
  \includegraphics[width=\linewidth]{trust-chain}
  \caption{Example of a trust-hierarchy. Here, the resolver can establish a trust chain between the Root and the ucla.edu zone.}
  \label{fig:trust-chain}
\end{figure*}

@@

\paragraph{Enabling DNSSEC} When enabling DNSSEC for a zone, the
parent zone has to establish and sign a DS record with the newly
created public-key.

\paragraph{Key Rollover} RFC4641 \cite{RFC4641} suggests that the DNSKEY
records for a zone should be changed somewhere between once every week
up to once every month. While the detailed rollover procedure is
irrelevant here, it is important to know that the parent zone has to
update its DS record regularly.

\paragraph{Disabling DNSSEC} An operator might choose to discontinue
DNSSEC for a zone. The DS records (or other means of signing DNSKEYs
as we will see later) need to be removed.

@@

This section will expose the current deployment status of DNSSEC and
give the reader an idea about the current DNSSEC landscape before
going into the specific DNSKEY management issues. Please refer to
section~\ref{sec:case-study} for more details on the situation of the
\verb|ch.| zone.

\subsection{Penetration in Production}
In figure~\ref{fig:dnssec-growth} you can see the development of
deployed, production DNSSEC zones. The data comes from SecSpider
\cite{secspider} which crawls as many DNSSEC zones as possible,
verifies whether they are properly signed and then applies a heuristic to
determine whether it is actually a production zone. The goal is to
rule out zones which are created for testing such as
\url{bogussig.bogussig.test.jelte.nlnet.labs.nl}.

We can see that DNSSEC is gradually being deployed and the number of
signed production zones grows rapidly (note the log-scale!). The rapid
increase of crawled DNSSEC zones mid-2008, so interprets
\cite{Osterweil09}, is due to a cache poisoning attack discovered in
summer 2008 \cite{dnsVuln}. During our research, we were unfortunately
unable to find an estimate of the number of (production) DNS zones in
the Internet to estimate the percentage of DNSSEC penetration.

@@

Another type of TAR resides at the resolver: A statically configured
TAR stores keys locally at the resolver and hence needs not to issue an
additional query for keys when a new DNS request arrives. The
statically configured TAR may automatically poll keys asynchronously,
or is maybe just a bunch of manually configured trusted keys. As an
example of a statically configured TAR, we have
Vantages\footnote{\url{http://www.vantage-points.org/}}, an addition
to a standard resolver, \cite{Osterweil09} which can poll keys from
different locations such as web pages, but also also other
vantage-enabled resolvers. Of course, the operator may also choose to
add trusted keys manually.

@@

The registries are the entities serving the records for a zone. For example, Switch is the registry for the \verb|ch.| zone. Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch.| is the school in itself (its IT-department to be precise \footnote{As a \emph{whois} query on the epfl.ch domain will show you.}), for example.

These two entities can eventually be linked by the third one, the registrar: this is where registrants will obtain the actual registration service, namely, the linking between a domain name and an IP address.

In Switzerland, for example, Switch plays both the roles of the registry and a registrar: other registrars also exist but one can purchase a domain name directly from Switch. Registries will generally delegate part or whole of the namespace registration managing to registrars\footnote{For example, one can also buy a .ch domain from \url{http://gandi.net}}

@@

The entities behind the three R's are the exact same whether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like IP addresses or billing data was actually updated.

@@

Enabling DNSSEC for a zone not being a \emph{do once and forget} operation, but instead requires additional coordination among the three R's, let us see what the managerial consequences arise, depending on if we cross or stay within certain administrative boundaries.

@@

\subsubsection{Multiple administrative domain}
DNSSEC's true challenges arise when we begin to consider several administrative entities (the three R's, mainly). As stated earlier, once the \verb|epfl.ch.| zone is signed it is easy for the school to sign \verb|www.epfl.ch.|. However, if the domain to be signed was \verb|epfl.ch.| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.| managed by Switch or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well be different among different registrars and TLD's.

While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant to such jitter: typically, if the \verb|epfl.ch.| zone updates its keys but Switch fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch.|'s DNS servers will appear as illegitimate whereas they perfectly are, and this will last as long as Switch's servers do not reflect the update.

As suggested by some RFC\cite{RFC4641} key updates should  happen at least every few months. Hence, it is crucial that a streamlined and reactive process is available for child-zones to notify their parent-zone when required. In the three R's setup, this means that the registry (or the registrars) must provide means for registrants to easily inform them of any changes regarding their DNSSEC keys.
%TODO process figure with Registrant -> Registrar -> Registry ?

@@

>>>>

\twocolumn[
\maketitle
\begin{onecolabstract}
  In this paper we recapitulate the current deployment status of
  DNSSEC and go into the main issues when managing its keys. This
  paper outlines the paper ``Interadministrative Challenges in
  Managing DNSKEYs'' \cite{Osterweil09} and updates some of its
  contents, especially with respect to the signing of the root zone on
  July 15, 2012. We will see that DNSSEC penetration in the Internet's 
  zones is still low, but that the necessary foundations have been lain
  for full deployment.
\end{onecolabstract}
\vspace{\baselineskip}
]

@@

In this paper we will first remind of the concepts and operations in
DNSSEC which are relevant for key management. Then we will give the
reader an overview of the current status of DNSSEC deployment in the
Internet to put the discussion about key management into the right
context.

Finally, we will shortly outline the DNSSEC situation in Switzerland
(i.e. the \verb|ch.| zone) and how its authorities handle DNSKEY management
issues.

@@

By signing the keys of the TLD's, the root will put its trust into their keys, and any keys the TLD's have signed, resulting in a trust hierarchy illustrated in figure~\ref{fig:trust-chain}. This way, a resolver querying the LACAL's DNS server for \verb|lacal.epfl.ch.| will be able to verify the received query answer against the known root-authority in the following way:

%% TODO this is not entirely true. The DNSKEYs are in the zone
%% itself. Only their signatures are in the parent zone.
\begin{enumerate}
\item Verify the reply's authenticity using \verb|lacal.epfl.ch.|'s public key, obtained from \verb|epfl.ch.|'s DNS server ;
\item Verify the query -- containing \verb|lacal.epfl.ch.|'s public key -- returned by \verb|epfl.ch.| with its public key, retrieved from the \verb|ch.| DNS server;
\item Again, verify \verb|ch.|'s responses by getting its public key from the root domain.
\item Finally, the root domain's responses can be verified with its already known public key.
\end{enumerate}

%% TODO you cannot talk about TARs here without explaining. Reference
%% for this?
DNSSEC's trust hierarchy is similar to the SSL Certificate infrastructure, where an entity marks its trust in another by \emph{signing} its certificate\footnote{Which is nothing more than a public key with additional metadata.}. However, DNSSEC theoretically only has one single top-authority\footnote{The use of TAR's should remain temporary, as in an ideal setup DNSSEC could work without them.}, while the SSL Certificate hierarchy disposes of many.

\begin{figure*}
\center
  \includegraphics[width=\linewidth]{trust-chain}
  \caption{Example of a trust-hierarchy. Here, the resolver can establish a trust chain between the Root and the \texttt{ucla.edu.} zone.}
  \label{fig:trust-chain}
\end{figure*}

@@

\paragraph{Enabling DNSSEC} When enabling DNSSEC for a zone, after it
has been signed, the parent zone has to establish and sign a DS record
with the newly created public-key.

\paragraph{Key Rollover} RFC4641 \cite{RFC4641} suggests that the DNSKEY
records for a zone should be changed somewhere between once every week
up to once every month. While the detailed rollover procedure is
irrelevant here, it is important to know that the parent zone has to
update its DS record regularly and hence has to be signed again, too.

\paragraph{Disabling DNSSEC} An operator might choose to discontinue
DNSSEC for a zone. The DS records in the parent zone (or other means
of signing DNSKEYs as we will see later) need to be removed.

@@

This section will expose the current deployment status of DNSSEC and
give the reader an idea about the current DNSSEC landscape before
going into the specific DNSKEY management issues. Please refer to
section~\ref{sec:case-study} for more details on the situation in the
\verb|ch.| zone.

\subsection{Penetration in Production}
In figure~\ref{fig:dnssec-growth} you can see the development of
deployed, production DNSSEC zones in time. The data comes from
SecSpider \cite{secspider} which crawls as many DNSSEC zones as
possible, verifies whether they are properly signed and then applies a
heuristic to determine whether it is actually a production zone. The
goal is to rule out zones which are created for testing such as
\url{bogussig.bogussig.test.jelte.nlnet.labs.nl}.

We can see that DNSSEC is gradually being deployed and the number of
signed production zones grows rapidly (note the log-scale!). The rapid
increase of crawled DNSSEC zones mid-2008, so interprets
\cite{Osterweil09}, is due to a cache poisoning attack discovered in
summer 2008 \cite{dnsVuln}. During our research, we were unfortunately
unable to find an estimate of the number of (production) DNS zones in
the Internet in order estimate the percentage of DNSSEC penetration.

@@

Another type of TAR resides at the resolver: A statically configured
TAR stores keys locally at the resolver and hence does not require to issue an
additional query for keys when a new DNS request arrives. The
statically configured TAR may automatically poll keys asynchronously,
or may be just a bunch of manually configured trusted keys. As an
example of a statically configured TAR, we have
Vantages\footnote{\url{http://www.vantage-points.org/}}, an addition
%% TODO what does this "plugin" do exactly?
to a standard resolver, \cite{Osterweil09} which can poll keys from
different locations such as web pages, but also also other
vantage-enabled resolvers. Of course, the operator may also choose to
add trusted keys manually.

@@

The registries are the entities serving the records for a zone. For example, SWITCH is the registry for the \verb|ch.| zone. Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch.| is the school in itself (its IT-department to be precise\footnote{As a \emph{whois} query on the epfl.ch domain will show you.}), for example.

%% TODO description about registrars is wrong. They do not host DNS entries.
These two entities can eventually be linked by the third one, the registrar: this is where registrants will obtain the actual registration service, namely, the linking between a domain name and an IP address. Figure~\ref{fig:3Rs} illustrates the relations among the three R's.

\begin{figure}
\center
  \includegraphics[width=\linewidth]{3Rs}
  \caption{Example of relations between the three R's.}
  \label{fig:3Rs}
\end{figure}

In Switzerland, for example, SWITCH plays both the roles of the registry and a registrar: other registrars also exist but one can purchase a domain name directly from SWITCH. Registries will generally delegate part or whole of the namespace registration managing to registrars\footnote{For example, one can also buy a .ch domain from \url{http://gandi.net}}

@@

%% TODO again: IP updates (except the Name Servers) don't go through 3R
The entities behind the three R's are the exact same whether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like IP addresses or billing data was actually updated.

@@

%% TODO phrase is a bit clumsy
Enabling DNSSEC for a zone not being a \emph{do once and forget} operation, but instead requiring additional coordination among the three R's, let us see what managerial consequences arise, depending on if we cross or stay within certain administrative boundaries.

@@

\subsubsection{Multiple administrative domain}
DNSSEC's true challenges arise when we begin to consider several administrative entities (the three R's, mainly). As stated earlier, once the \verb|epfl.ch.| zone is signed it is easy for the school to sign \verb|www.epfl.ch.|. However, if the domain to be signed was \verb|epfl.ch.| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.| managed by SWITCH or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well be different among different registrars and TLD's.

While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant to such jitter: typically, if the \verb|epfl.ch.| zone updates its keys but SWITCH fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch.|'s DNS servers will appear as illegitimate whereas they perfectly are, and this will last as long as SWITCH's servers do not reflect the update.

As suggested by RFC4641 \cite{RFC4641} key updates should  happen at least every few months. Hence, it is crucial that a streamlined and reactive process is available for child-zones to notify their parent-zone when required. In the three R's setup, this means that the registry (or the registrars) must provide means for registrants to easily inform them of any changes regarding their DNSSEC keys.
%TODO process figure with Registrant -> Registrar -> Registry ?

@@

%% TODO add part about computation
